Analytics Without Parameter Tuning Considered Harmful?

### 2015-03-13 learners.tex
Updates from ShareLaTeX

This section describes the learners used in this study. While there are many differences
in those learners, they have the same goal: to summarize a large data set into a small model\footnote{Technical note: proponents of instance-based reasoning might argue that
their goal is to retain, and not summarize, the training data.
However, 
even instance-based reasoners use some degree of summarizing when they learn
ways to bias the instance matching away from irrelevant training data.
Those bias tools include  generalizing points to ranges (to allow for partial matches)
or  clustering algorithm that restrict the case matching
to just the subset of old cases most similar  to new  cases.}.

For example,
consider some learner that
builds predictive rules from a subset of  attributes $A$ ranges. Assuming numeric attributes
divided into $d=5$ divisions, then the 20 attributes of \fig{ck} can be combined into 
$2^{A*d} = 2^{5*20} > 10^{30}$
possible rules. The key thing to note here is that many of these rules will have similar effects
since, at least in software engineering, data sets containing rows $R$ and columns $C$ can contain much redundancy:
\bi
\item When $c \subset C$ columns of data that are correlated to each other, then those
$c$ columns can be replaced by one column.
\item When $r \subset R$ rows of data are very similar (e.g. they come from similar projects), 
then those $r$ rows can be replaced
by one cluster centroid that stands in $r$.
\item
Kocagenunil~\cite{me13a} and Papakroni~\cite{papa13} conduct  different experiments
to  show  that tables of software defect data in the form of \fig{ck} contain around 90\% redundant
data. Specifically, they show that such tables can be be pruned back to around
25\% of the columns and 10\% of the rows without losing the essential information in the data set.
\ei
Given this high level of redundancy, learners need only explore a small fraction of these $2^{A*d}$ rules.
Most learners exploit this fact to optimize their reasoning. Rather than explore all options,
they employ some {\em search bias} that lets them quickly focus on a small set of promising rules.

>>>>

This section describes the learners used in this study: CART~\cite{brieman00}, Random Forest~\cite{breiman84}, 
and WHERE~\cite{menzies2013local}. The implementations
for CART and Random Forest comes from 
SciKitLearn~\cite{scikit-learn}.
WHERE is available from
github.com/ai-se/where\footnote{\wei{ need a nice  where repo. with an
data and code examples sub-directory. clean code. throw out anything not needed. before april1.}}

CART, Random Forest, and WHERE are all  tree learners that divide a data set, then recurses
on each split.
All our learners use the following tuning parameters to control that splitting process.
These learners
generate numeric predictions which are converted
into binary ``yes/no'' decisions via \eq{yesno}. Hence, they all use the threshold value $T$ discussed in \tion{eq}.
Also, if data contains more than {\em min sample split}, then a split is attempted.
On the other hand, if a split contains no more than {\em min samples leaf}, then recursion stops. For CART and Random Forest use a 
user-supplied constant for this parameter while
WHERE computes $m$={\em min samples leaf} from the size of the data
sets via  $m=\mathit{size}^\mathif{min size}$ (so, for WHERE,
{\em min size} is the parameter to be tuned).

These learners use different techniques to explore the splits:
\bi
\item
CART finds the attributes whose ranges contain rows with least variance in the number
of defects\footnote{If an attribute ranges $r_i$ is found in 
$n_i$ rows each with a  defect count variance of $v_i$, then CART seeks the attributes
whose ranges minimizes $\sum_i \left(\sqrt{v_i}\times n_i/(\sum_i n_i)\right)$.}.
\item
Random Forest    divides data like CART,
but it builds more than one tree, each time using some subset of
the attributes (selected at random). 
\item
WHERE projects the data on to dimensions synthesizes from the raw data using
a process analogous to principle component analysis\footnote{
PCA  synthesises  new
attributes $e_i, e_2,...$
that extends across the dimension of greatest  variance in the data  with attributes $d$.  
This process  combines
redundant  variables into a smaller set of variables  (so $e \ll d$) since those
redundancies become (approximately) parallel lines
in $e$ space. For all such redundancies \mbox{$i,j \in d$}, we 
can ignore $j$ 
since effects that change over $j$ also
change in the same way over $i$.
PCA is also useful for skipping over noisy variables from $d$-- these
variables are effectively ignored since    they  do not contribute to the variance in the data.}.
WHERE   divides  at the median point of that projection. On recursion,
this generates a dendogram, the leaves of which are clusters of  very similar examples.
\ei
WHERE's {\em infoPrune} tuning parameter then uses  what attributes are used in this process.

XXX

Once WHERE's dengogram is generated, some mechanism must be used to map test cases into specific 
clusters. To this end, WHERE pretends its clusters are ``classes'' and 
a entropy-based decision tree learner   to find
the decisions that best select for each ``class'' (i.e. each cluster).
This entropy-based learner  finds the attributes that build sub-trees that best select
for the fewest  clusters\footnote{In entropy based decision-tree learning,  the
the splitting attribute is the one that most  minimizes $\sum_i \left( e_i\times n_i/(\sum_i n_i)\right)$,
defined as follows.
Let    attribute ranges $r_i$ select
$n_i$ rows that are found in $c_j$ clusters.  Then the probability of
arriving at those clusters is $p_j = c_j/n_i$ and the entropy
of those rows is $e_i = \sum_j \left( -p_j \ldot log_2p_j\right)$.}.
\ei
After b

If WHERE's {\em tree prune} parameter is enabled, then WHERE also prunes  superfluous sub-trees
in the cluster-selection trees. For example, if a sub-tree and its parent have the same 
majority cluster
(one that occurs most frequently), then we prune the sub-tree.

Other tuning parameters are learner specific. For example,
{\em max feature} is used by
CART and Random Forest to select the number of attributes
used to build one tree.
CART's default is to use all the attributes while 
Random Forest usually selects the square root of the number
of attributes.
Also,
$F$ is the number of trees built by Random Forest 
(and was
  discussed in our introduction).
 Further,
  {\em max leaf nodes} is the upper bound on leaf notes generated in a 
  Random Forest.

Of all our learners, WHERE has the most tuning parameters. This turns out
to be significant since, in the experiments shown below, tuned WHERE 
out-performed the other learners. This suggests the heuristic that if
you are going to tune learners, use one with the most tuning options.

After building its tree, WHERE ranks all attributes by how well
they select for fewest leaves (using the  infogain measure
from the Fayyad-Irani discretizer~\cite{FayIra93Multi}, where ``class'' is really the leaf cluster ids). It then only uses the best {\em infoPrune} percent of the attributes to build
its cluster selection tree. Also, the {\em tree prune} parameter (described above) sometimes
prunes aways all  cluster selectors branches. To tame this effect, the {\em wriggle} parameter
blocks tree pruning for at least the the   {\em wriggle} number of initial branches.

We choose CART and Random Forest since  these were used in 
a recent IEEE TSE paper by Lessmann et al.~\cite{lessmann2008benchmarking} that compared 21 different 
learners for software defect prediction:
\bi
\item
{\em Statistical classifiers:}
Linear    discriminant analysis,
Quadratic discriminant analysis,
Logistic regression,
Naive Bayes,
Bayesian networks,
Least-angle regression,
Relevance vector machine,

\item
{\em Nearest neighbor methods:}
k-nearest neighbor,
K-Star

\item
{\em Neural networks:}
Multi-Layer Perceptron,
Radial bias function network,

\item
{\em Support vector machine-based classifiers:}
Support vector machine,
Lagrangian SVM
Least squares SVM,
Linear programming,
Voted perceptron,

\item
{\em Decision-tree approaches:}
C4.5 decision tree,
CART,
Alternating decision tree.
\item
{\em Ensemble methods:}
Random Forest,
Logistic Model Tree.
\ei
In that study, CART was severely trounced (was ranked last) and Random Forest was
the standout best method. The experiments shown below both confirm and refute
that ranking. In a result consistent with the prior result, untuned Random Forest performs best.
However, after tuning, the worst learner found by Lessmann et al. (CART) performed better
than Random Forest.

, where the performance of both tools as predictors are 
significantly different based on authors' experiment as mentioned before. We'd like to 
investigate whether tuning can change such conclusion. WHERE-based Predictor is a new 
defect predictor based on WHERE\cite{menzies2013local} algorithm. A comparison with 
standard predictor like CART and Random Forest will better evaluate and judge the 
performance such new predictor. As for the tuner, there're many heuristic optimization 
algorithms in wild. However, DE is a good but maybe not the best candidate for the tuning 
process considering different performance measurements. To determine which optimizer is 
fitable and results in better performance is beyond this work scope. We leave it to future 
works. The rest of this section will describe each tool applied in this work.

While there are many technical
differences
in those learners, they have certain similarities.

To begin with, all the learners in this study generate numeric predictions which are converted
into binary ``yes/no'' decisions via \eq{yesno}. Hence, they all have some threshold  value $T$ 
that can be tuned.

same goal: to summarize a large data set into a small model\footnote{Technical note: proponents of instance-based reasoning might argue that
their goal is to retain, and not summarize, the training data.
However, 
even instance-based reasoners use some degree of summarizing when they learn
ways to bias the instance matching away from irrelevant training data.
Those bias tools include  generalizing points to ranges (to allow for partial matches)
or  clustering algorithm that restrict the case matching
to just the subset of old cases most similar  to new  cases.}.

For example,
consider some learner that
builds predictive rules from a subset of  attributes $A$ ranges. Assuming numeric attributes
divided into $d=5$ divisions, then the 20 attributes of \fig{ck} can be combined into 
$2^{A*d} = 2^{5*20} > 10^{30}$
possible rules. The key thing to note here is that many of these rules will have similar effects
since, at least in software engineering, data sets containing rows $R$ and columns $C$ can contain much redundancy:
\bi
\item When $c \subset C$ columns of data that are correlated to each other, then those
$c$ columns can be replaced by one column.
\item When $r \subset R$ rows of data are very similar (e.g. they come from similar projects), 
then those $r$ rows can be replaced
by one cluster centroid that stands in $r$.
\item
The seperate experiments of
Kocagenunil~\cite{me13a} and Papakroni~\cite{papa13}  show that  tables  of data
like \fig{ck} can be be pruned back to around
25\% of the columns and 10\% of the rows without losing the essential information in the data set.
This reduced space contains $0.25*0.1 = 2.5\%$ of the cells in the original table.
While some of that reduction is due to the removal of noise, their conclusion is correlated
columns and similar rows are very common in SE data sets.
\ei
Given this high level of redundancy, learners need only explore a small fraction of these $2^{A*d}$ rules.
Most learners exploit this fact to optimize their reasoning. Rather than explore all options,
they employ some {\em search bias} that lets them quickly focus on a small set of promising rules.

### 2015-03-13 tuningpredictor.tex
Updates from ShareLaTeX

@@

Thirdly, it is standard practice to use the default ``off-the-shelf'' tunings  for data mining tools (previously
we have defended that approach on methodological grounds arguing that it
encourages reproducibility~\cite{me15:book1}). That ``off-the-shelf''  policy
can no longer be condoned. For example, one such default setting
  in the Python SciKitLearn toolkit~\cite{scikit-learn}
is to use $T=10$ decision trees in  Random Forest classifiers.
Our optimizer settled on  $T$ values that were were nowhere near that default:

{\scriptsize
\[T \in \left\{\begin{array}{l} 55,  65, 70,   82, 88, 96, 100,  102,  104, 107,\\
                                108,  119, 133,  140, 140,   147,  145,  142   \end{array}\right\}
\]}

In summary,it is now an open and pressing research issue to check if
analytics without parameter tuning is considered {\em harmful} or, at the 
very least, {\em misleading}.
At the very least, there is now cause to doubt many conclusions based on
``off-the-shelf'' tunings.
At the very most, 
it is no longer enough to just run a data miner and present the result
{\em without} first conducting an tuning optimizations study.

The rest of this paper is structured as follows.

\section{Preliminaries}

Before beginning, this section offers certain caveats on these results.

@@

Secondly, the tuning results shown here only came from one particular software analytics task 
(defect prediction from static code attributes).
There are many other kinds of software analytics tasks 
(software development effort estimation, social network mining,
detecting duplicate issue reports, etc) and the implication of this
study for those tasks is unclear. 
However,  those other tasks often use the same kinds of learners
explored in this paper so it is quite possible that
the conclusions of this paper apply to other SE analytics tasks as well.

@@

\section{Background}
\subsection{Defect Prediction}

@@

\subsection{Data Mining}
Static code defect prediction applies data miners to build its predictors. How does data mining work?

>>>>

\newcommand{\wei}[1]{\textcolor{red}{Wei: #1}}

@@

Thirdly, it is standard practice to use the default ``off-the-shelf'' tunings  for data mining tools (previously
we have defended that approach on methodological grounds arguing that it
encourages reproducibility~\cite{me15:book1}). That ``off-the-shelf''  policy
can no longer be condoned. For example, one such default setting
  in the Python \mbox{SciKitLearn} toolkit~\cite{scikit-learn}
is to use $F=10$ decision trees in  Random Forest classifiers.
Our optimizer settled on  $F$ values that were were nowhere near that default:

{\scriptsize
\[F \in \left\{\begin{array}{l} 55,  65, 70,   82, 88, 96, 100,  102,  104, 107,\\
                                108,  119, 133,  140, 140,   147,  145,  142   \end{array}\right\}
\]}

In summary,it is now an open and pressing research issue to check if
analytics without parameter tuning is considered {\em harmful} or, at the 
very least, {\em misleading}.
Clearly, we must now doubt  conclusions based on
``off-the-shelf'' tunings.
Further,
it is no longer enough to just run a data miner and report the result
{\em without} first conducting an tuning optimizations study.

\subsection{Preliminaries}

Before beginning, we offer four caveats on these results.

@@

Secondly, the tuning results shown here only came from one  software analytics task 
(defect prediction from static code attributes).
There are many other kinds of software analytics tasks 
(software development effort estimation, social network mining,
detecting duplicate issue reports, etc) and the implication of this
study for those tasks is unclear. 
However,  those other tasks often use the same kinds of learners
explored in this paper so it is quite possible that
the conclusions of this paper apply to other SE analytics tasks as well.

@@

\section{Motivating Example}\label{sec:eg}

This section offers a small demonstration
of the impact of tuning parameters. It also demonstrates that this issue of tuning effects not jist the complex data miners discussed later in the paper bu t also 
applies for even very simple  learning schemes.

One way to generate a  defect predictor from
static code is to use 
linear  regression.
This is a standard statistical
method that fits a straight line to a set of points. The line
offers a set of predicted values.
If the points are somewhat
scattered, then a single regression line cannot pass through all points and the distance from these predicted values to the actual
values is a measure of the error associated with that line.
Linear regression search for a line that minimizes that
error and maximizes the correlation\footnote{
$Correlation$
measures how closely two variables co-vary.  It ranges from   from -1
(perfect negative correlation) through 0 (no correlation) to +1
 (perfect positive correlation).
Let  $a_i$ and $p_i$ denote some actual and predicted values respectively; and  $n$ and $\overline{x}$ denote
 the number of observations and the mean of the $n$ observations, 
respectively. Then:
$S_{PA}=(\sum_i (p_i - \overline{p})(a_i -
  \overline{a}))/(n-1)$ and
$S_{p}=(\sum_i (p_i - \overline{p})^2)(n-1)$
and 
$S_{a}=(\sum_i (a_i - \overline{a})^2)/(n-1)$ and
correlation is  $c=S_{PA}/\sqrt{S_pS_a}$.}  denoted as $c$.

Suppose  a researcher wants to use linear regression
to test if Halstead's~\cite{halstead77} measures
of   function complexity
(number of symbols programmers has to understand) are   {\em better than}
mere lines of code for predicting
software defects.  That researcher might argue that Halstead's cognitive approach to
software bugs is better suited to code refactoring tools since it offers 
more ways to alter functions that just some coarse grain lines of code measure.

That researcher might test that belief by using studying historical
defect logs with 
linear regression. Here are two equations (learned from the NASA data at goo.gl/pGDfvp)
that use just lines of code or the Halstead measures $N,V,L,D,I,E,B,T$ seen in a
software module (in this case, a ``C'' function).
Note that the Halstead correlation $c_2$
is worse than those from lines of code $c_1$. This result  seems to suggest that
despite their potential support for refactoring, our researchers should not use Halstead.

{\scriptsize \[\begin{array}{l|l|ll}
\mathit{measures} & d= \mathit{\#defects} & \mathit{correlation}\\\hline
\mathit{LOC}   &d_i= 0.0164 +0.0114\mathit{LOC}\ & c_1 = 0.65\\\hline
\mathid{Halstead & d_2= 0.231 + 0.00344N     \\    
           & +  0.0009V   - 0.185L \\
           & - 0.0343D      - 0.00541I \\
           & + 0.000019E + 0.711B \\ 
            &   - 0.00047T  & c_2=-0.36 
            \end{array}
\end{eqnarray}}

\noindent
Enter tuning. Suppose the defect predictors $d_1$ and $d_2$  learned from LOC or Halstead
are used to to call an inspection
team to check for errors in certain parts of the code using the rule
\begin{equation}\label{eq:yesno}\scriptsize
\mathit{inspect}= \left\{
\begin{array}{ll}
d_i \ge T \rightarrow \mathit{Yes}\\
d_i <   T \rightarrow \mathit{No}
\end{array}
\end{equation}
\fig{pd1} shows the effects of tuning the $T$ variable. Not surprisingly,
at $T=0$, all modules get inspected so the false alarm rate is very high. To reduce that
problem, we can increase $T$. \fig{pd1} reports that the false alarm rate falls below
20\% at $T=0.45$ (for Halstead).

\begin{figure}[!t]
{\scriptsize
\begin{center}

\% recall (probability of detection):

\includegraphics[width=3in]{lsrvscostpd.pdf}

\% false alarms:

\includegraphics[width=3in]{lsrvscostpf.pdf}
\end{center}}
\caption{
 Y-axis shows probability of false alarm,
  and
  probability of recognizing defective modules  seen using \mbox{$d_i \ge T$}.
  Curves calculated from the KC2 dataset from the PROMISE repository goo.gl/pGDfvp.
 }\label{fig:pd1}
 \end{figure}

One  important lesson from  \fig{pd1} is that the ``best'' tunings are context
specific. For mission critical systems (e.g. a nuclear power plant), management
might accept the cost of high false alarm rates if it meant increasing the probability
of detecting errors. For such contexts, we might recommend some very small value of $T$.

The other  important lesson from   \fig{pd1}  is that, with tuning, a seemingly poor
detector can work just as well as seemingly better ones.
Note that either the Halstead or LOC detector can reach some desired
level of recall, regardless of their correlations, just by
selecting the appropriate threshold value. For example, in \fig{pd1}, see the recall=75\% values
found at {\em either} $d_i\ge 0.65$ or $d_2\ge 0.45$ (and at the threshold, the false alarm rates
were very similar: 14\% and 19\%).

The core point of this example is that    the true value of a detector
could not be assessed {\em without} conducting a  tuning study in the context of some business case (in this case, 
issuing a request to an inspection team to review some module).  
This is strong motivation to explore the issue of tunings.

\section{Background Notes}

The rest of this paper repeats the analyis of the last section, but for much
more complex learners. In those results, shown  below, we will find
examples of {\em rank reversals};
i.e.  after tuning the top ranked defector becomes last and vice versa.
Such examples motivate the extensive exploration of tuning in software analytics.
Before presenting those results, this section offers some background notes on defect prediction,
performance measures, and optimization with differential evolution.

\subsection{Defect Prediction}

@@

\subsection{Data Mining Algorithms}
